Jane, John … Leslie? A Historical Method for Algorithmic Gender Detection

### 2014-09-18 article.Rmd
Color Palette

```{r load-libraries, include=FALSE}
options(stringsAsFactors = FALSE)
library(genderdata)
library(gender)
library(dplyr)
library(ggplot2)
library(mullenMisc)
library(knitr)
library(stringr)
library(scales)
library(tidyr)

@@

article_theme <- theme(legend.title = element_blank())
```

@@

```{r}
diss_predicted %>%
  ggplot(aes(x = year, fill = gender_author)) + 
  geom_bar(binwidth = 1) +
  ggtitle("Dissertations about History by Gender, 1950-2012") +
  xlab(NULL) + ylab(NULL) +
  short_scale + theme(legend.position = "bottom") +
  scale_fill_discrete(name = "",
                      labels = c("Female", "Male")) +
  article_theme
```

@@

diss_proportions  %>%
  ggplot(aes(x = year, y = proportion, fill = gender)) +
  geom_bar(binwidth = 1, stat = "identity") +
  ggtitle("Proportion of Dissertations about History by Gender, 1950-2012") +
  xlab(NULL) + ylab(NULL) +
  short_scale + theme(legend.position = "bottom") +
  geom_hline(yintercept = 0.5) +
  scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
  article_theme
```

@@

ahr_proportions %>%
  ggplot(aes(x = year, y = proportion, fill = gender)) + 
  geom_bar(binwidth = 1, stat = "identity") +
  ggtitle("Authors of Books Reviewed in the AHR by Gender") +
  xlab(NULL) + ylab(NULL) +
  theme(legend.position = "bottom") +
  geom_hline(yintercept = 0.5) +
  scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
  article_theme
```

@@

combined_proportions %>%
  filter(gender != "unknown_proportion") %>%
  ggplot(aes(x = year, y = proportion, color = gender, size = type)) + 
  geom_line() +
  short_scale +
  geom_hline(yintercept = 0.5) +
  scale_size_discrete(name = "Type",
                      labels = c("Book", "Dissertation"), range = c(2,1)) +
  scale_color_discrete(name = "Gender",
                       labels = c("Female", "Male")) +
  xlab(NULL) + ylab("Proportion of authors by gender") +
  ggtitle("The gap between the genders of book and dissertation authors")
```

@@

![Gender of reviewers of books authored by women](img/female_authors_withreviewers.jpg)

@@

![Gender of reviewers by category of review](img/categories_gender_bytime.jpg)

>>>>

```{r load-libraries, include=FALSE}
options(stringsAsFactors = FALSE)
library(genderdata)
library(gender)
library(dplyr)
library(ggplot2)
library(mullenMisc)
library(knitr)
library(stringr)
library(scales)
library(tidyr)
library(ggthemes)
library(RColorBrewer)

@@

article_theme <- theme(legend.title = element_blank())

article_palette <- "Dark2"
```

@@

```{r}
diss_predicted %>%
  ggplot(aes(x = year, fill = gender_author)) + 
  geom_bar(binwidth = 1) +
  ggtitle("Dissertations about History by Gender, 1950-2012") +
  xlab(NULL) + ylab(NULL) +
  short_scale + theme(legend.position = "bottom") +
  scale_fill_brewer(palette = article_palette, name = "",
                      labels = c("Female", "Male")) +
  article_theme
```

@@

diss_proportions  %>%
  ggplot(aes(x = year, y = proportion, fill = gender)) +
  geom_bar(binwidth = 1, stat = "identity") +
  ggtitle("Proportion of Dissertations about History by Gender, 1950-2012") +
  xlab(NULL) + ylab(NULL) +
  short_scale + theme(legend.position = "bottom") +
  geom_hline(yintercept = 0.5) +
  scale_fill_brewer(palette = article_palette,
                    labels = c("Female", "Male", "Unknown")) +
  article_theme
```

@@

ahr_proportions %>%
  ggplot(aes(x = year, y = proportion, fill = gender)) + 
  geom_bar(binwidth = 1, stat = "identity") +
  ggtitle("Authors of Books Reviewed in the AHR by Gender") +
  xlab(NULL) + ylab(NULL) +
  theme(legend.position = "bottom") +
  geom_hline(yintercept = 0.5) +
  scale_fill_brewer(palette = article_palette,
                      labels = c("Female", "Male", "Unknown")) +
  article_theme
```

@@

combined_proportions %>%
  filter(gender != "unknown_proportion") %>%
  ggplot(aes(x = year, y = proportion, color = gender, size = type)) + 
  geom_line() +
  short_scale +
  scale_size_discrete(name = "Type",
                      labels = c("Book", "Dissertation"), range = c(2,1)) +
  scale_color_brewer(palette = article_palette, name = "Gender",
                       labels = c("Female", "Male")) +
  xlab(NULL) + ylab("Proportion of authors by gender") +
  ggtitle("The gap between the genders of book and dissertation authors")
```

@@

```{r}
ahr_predicted %>%
  filter(gender_author == "female") %>%
  select(year, gender_reviewer) %>%  
  group_by(year, gender_reviewer) %>% 
  summarize(n = n()) %>% 
  spread(gender_reviewer, n, fill = 0) %>%
  mutate(female_proportion = female / (female + male + unknown),
         male_proportion = male / (female + male + unknown),
         unknown_proportion = unknown / (female + male + unknown)) %>%  
  select(female_proportion, male_proportion, unknown_proportion, year) %>%
  gather(gender, proportion, -year) %>%
  ggplot(aes(x = year, y = proportion, fill = gender)) + 
  geom_bar(binwidth = 1, stat = "identity") +
  ggtitle("Gender of reviewers of female book authors in AHR") +
  xlab(NULL) + ylab(NULL) +
  theme(legend.position = "bottom") +
  geom_hline(yintercept = 0.5) +
  scale_fill_brewer(palette = article_palette, 
                    labels = c("Female", "Male", "Unknown")) +
  article_theme +
  short_scale
```

@@

**Cameron: Do you have the mapping you use from the AHR's 68 categories to the 8 or so you used? With that, I just have one or two more lines to facet this plot. (It doesn't look right because it's not faceted yet.)**

```{r}
ahr_predicted %>%
  filter(gender_author == "female") %>%
  select(year, gender_reviewer, category) %>% 
  group_by(year, gender_reviewer, category) %>%  
  summarize(n = n()) %>%
  spread(gender_reviewer, n, fill = 0) %>% 
  mutate(female_proportion = female / (female + male + unknown),
         male_proportion = male / (female + male + unknown),
         unknown_proportion = unknown / (female + male + unknown)) %>%  
  select(female_proportion, male_proportion, unknown_proportion,
         year, category) %>% 
  gather(gender, proportion, -year, -category) %>%
ggplot(aes(x = year, y = proportion, fill = gender)) + 
  geom_bar(binwidth = 1, stat = "identity") +
  ggtitle("Gender of reviewers of female book authors in AHR") +
  xlab(NULL) + ylab(NULL) +
  theme(legend.position = "bottom") +
  geom_hline(yintercept = 0.5) +
  scale_fill_brewer(palette = article_palette,
                    labels = c("Female", "Male", "Unknown")) +
  article_theme +
  short_scale
```
